\textbf{Transparency to hybrid open access through publisher-provided metadata: An article-level study of Elsevier}

### 2021-05-17 analysis/manuscript.Rmd
revise geographical distribution on central invoicing receipients

```{r}
# load eligible oa articles
hybrid_volume <- readr::read_csv(here::here("data", "hybrid_articles.csv"))

# aggregate data
sponsored_articles <- hybrid_volume %>%
  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND")) %>%
  count(issued_year, fct_source, cc_norm)
all_articles <- hybrid_volume %>%
  group_by(issued_year) %>%
  summarise(n = n())
sponsored_art <- sponsored_articles %>% 
  group_by(fct_source) %>% 
  summarise(nn = sum(n)) %>% 
  #mutate(cumsum = cumsum(nn)) %>% 
  mutate(prop = nn / sum(nn) * 100) %>%
 # mutate(cumprop = cumsum(prop)) %>%
  adorn_totals("row")
```

@@

```{r license_count}
sponsor_df <- hybrid_invoice %>% 
#  filter(!is.na(Country)) %>% 
  # manual changes
  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "European Union", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf))
```

>>>>

```{r}
# load eligible oa articles
hybrid_volume <- readr::read_csv(here::here("data", "hybrid_articles.csv")) 
# add open apc info
hybrid_volume <- hybrid_volume %>%
      mutate(in_open_apc = tolower(doi) %in% tolower(o_apc$doi))
# aggregate data
sponsored_articles <- hybrid_volume %>%
  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC BY", "CC BY-NC-ND")) %>%
  count(issued_year, fct_source, cc_norm)
all_articles <- hybrid_volume %>%
  group_by(issued_year) %>%
  summarise(n = n())
sponsored_art <- sponsored_articles %>% 
  group_by(fct_source) %>% 
  summarise(nn = sum(n)) %>% 
  #mutate(cumsum = cumsum(nn)) %>% 
  mutate(prop = nn / sum(nn) * 100)  %>%
 # mutate(cumprop = cumsum(prop)) %>%
  adorn_totals("row")
```

@@

```{r license_count}
sponsor_df <- hybrid_invoice %>% 
#  filter(!is.na(Country)) %>% 
  # manual changes
  mutate(Country = ifelse(grepl("EU", Country), "European Union", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf))
```

### 2021-05-17 analysis/manuscript.tex
revise geographical distribution on central invoicing receipients

Elsevier's data offer more insight into invoicing. Overall, we
identified 62 sponsoring institutions and organizations that received
invoices for hybrid publishing fees as part of an agreement with
Elsevier. We found that, by a large margin, most invoices were issued to
UK-based research funders and institutions (n=14,344; 59.2\%), followed
by the Netherlands (n=2,835; 11.7\%) and the European Union (n=2,164;
8.9\%; EU). Figure \ref{fig:invoice_sponsor_country} provides a closer
look at the geographical distribution by license prevalence,
highlighting the dominance of CC -BY licensed articles invoiced to
sponsoring institutions in the UK, the United States, and, to a much
lesser extent, the Netherlands. In contrast, the majority of articles
invoiced to sponsoring institutions from Norway or representing the EU
were published under a non-commercial license.

>>>>

Elsevier's data offer more insight into invoicing. Overall, we
identified 62 sponsoring institutions and organizations that received
invoices for hybrid publishing fees as part of an agreement with
Elsevier. We found that, by a large margin, most invoices were issued to
UK-based research funders and institutions (n=14,344; 59.2\%), followed
by the Netherlands (n=2,835; 11.7\%) and the European Union (n=1,990;
8.2\%; EU). Figure \ref{fig:invoice_sponsor_country} provides a closer
look at the geographical distribution by license prevalence,
highlighting the dominance of CC -BY licensed articles invoiced to
sponsoring institutions in the UK, the United States, and, to a much
lesser extent, the Netherlands. In contrast, the majority of articles
invoiced to sponsoring institutions from Norway or representing the EU
were published under a non-commercial license.
